---
title: "Sat test Data"
author: "Jenny Xiao"
date: "10/15/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    df_print: paged
    code_folding: hide
    toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

## Info {.tabset}

### Import a XML file into R
To deal with xml file, we need use XML or xml2 packages.  xml2 has similar goals to the XML package. The main differences are:
* xml2 takes care of memory management for you. It will automatically free the memory used by an XML document as soon as the last reference to it goes away.

* xml2 has a very simple class hierarchy so don't need to think about exactly what type of object you have, xml2 will just do the right thing.

* More convenient handling of namespaces in Xpath expressions - see xml_ns() and xml_ns_strip() to get started.

I picked XML package and it seems working well.

    
### Introduce the Weather data set
[Here you can get a link](https://catalog.data.gov/dataset/local-weather-archive) for downloading the xml data.
This dataset contains data by school on number of student who participate SAT,participate rate and the percent meeting benchmark of 2012 and 2013.
There are not categorial variables in this data set. I add two columns of categorial data according the the test_takers_change and the participation_rate_estimate_change

## Data Loading 
Here is part of the dataset.

```{r results='hide', message=FALSE, warning=FALSE}

library("XML")
library("methods")
library(dplyr)
library(kableExtra)
library(tidyverse)


xmlfile<-xmlParse(file = "satdata.xml")
my_data<-xmlToDataFrame(node=getNodeSet(xmlfile,"//response/row/row"))

kable(my_data, "html", caption="SAT School Participation and Performance: 2012-2013") %>%
  kable_styling(bootstrap_options = c( "hover"))

```

### Categorical summeries and Granph analysis 
We can see from the plot,for the schools which test_takers rised the partipate_rate_change also rised,no dropps.

```{r}
# Add two category variable to my dataset
cate_data<-my_data %>% 
  mutate(test_t_c=ifelse(my_data$test_takers_change < 0, "Dropped","Rised"))

cate_data<- cate_data %>% mutate(part_r_c=ifelse(my_data$participation_rate_estimate_change < 0, "Dropped","Rised")) %>%
  drop_na(part_r_c) %>% drop_na(test_t_c)

ggplot(cate_data, aes(x = test_t_c))+ geom_bar() +
  labs(x = "Test takes changing plot for each school",title = "Comparision on 2013 and 2012 the test takers changing in terms of each school")


ggplot(cate_data, aes(x = test_t_c)) + geom_bar(aes(fill=part_r_c),position = "dodge") +
    labs(x = "Test takers Change",title = "The relation between Rain and wind direction")

```

### Numeric Summaries for some schools

  Here i just picked four variables to do the numeric summary analysis.
 

```{r}
get_numericSummary<-function(x, ...){
  return(list(round(min(x, na.rm = TRUE),digits=1),
              round(quantile(x,0.25, na.rm=TRUE,names = FALSE),digits = 1), 
              round(quantile(x,0.5,na.rm=TRUE,names = FALSE),digits = 1), 
              round(mean(x, na.rm = TRUE), digits=1), 
              round(quantile(x,0.75,na.rm=TRUE,names = FALSE), digits = 1),
              round(max(x, digits=1,na.rm = TRUE),digits = 1)))
  
}
my_data$test_takers_2012 <- as.numeric(as.character(my_data$test_takers_2012))
my_data$test_takers_2013 <- as.numeric(as.character(my_data$test_takers_2013))
my_data$percent_meeting_benchmark_2012 <- as.numeric(as.character(my_data$percent_meeting_benchmark_2012))
my_data$percent_meeting_benchmark_2013 <- as.numeric(as.character(my_data$percent_meeting_benchmark_2013))

a <-get_numericSummary(my_data$test_takers_2012)
b <-get_numericSummary(my_data$test_takers_2013)
c <-get_numericSummary(my_data$percent_meeting_benchmark_2012)
d <-get_numericSummary(my_data$percent_meeting_benchmark_2013)

summary_table <- cbind(a,b,c,d)
rownames(summary_table) <-c("Min", "1st Qu", "Median", "Mean", "3rd Qu", "Max")
colnames(summary_table) <-c("test_takers_2012","test_takers_2013","percent_meeting_benchmark_2012","percent_meeting_benchmark_2013")
kable(summary_table, "html", caption="Summary of Each school") %>%
  kable_styling(bootstrap_options = c( "hover"))


```

### Scatterplot

Here we use scatterplot show the correlation and regression betwween the test_takers_2012 and test_takers_2012, and also the percent_meeting_benchmark_2013 and percent_meeting_benchmark_2013.

```{r message=FALSE,warning=FALSE}

g <- ggplot(my_data, aes(x = test_takers_2012, y = test_takers_2013))+
  geom_point( method= lm)
print(g + ggtitle("2012 test takers capare to the 2013 test takers ")) 

g <- ggplot(my_data, aes(x = percent_meeting_benchmark_2013, y = percent_meeting_benchmark_2012, ))+
  geom_point() 
print(g + ggtitle("2013 the pecent meeting benchmark compare to the 2012")) 

```



